Supporting Information : Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-11-19 Ben_Manuscripts/structure_paper/supporting_information.tex
Complete draft of response to reviewers

  \bibliographystyle{ieeetr}
  \graphicspath{{./figures/supporting_figures/}}  % put all the figures here
  \maketitle

@@

@@

@@

  \section{Correlation Functions}

@@

@@

  The appearance of R-double is likely a consequence of structure induced when
  vertically stacked monomer pairs share hydrogen bonds with a common water molecule (See
  Section~\ref{M-section:rdouble} of the main text). We quantified the occurence of
  these pairing interactions by binning the $z$-positions of the center of mass of 
  the monomer head groups into 20 bins, the same number as the monomers per column. 
  This works relatively well since there is a high degree of correlation between
  columns in our system. Peaks in the distributions represent the scenario where a monomer head group
  shares a water molecule with a head group vertically above it (See Figure~\ref{fig:pore_hbonds}).
  We calculated the discrete Fourier transform of the distributions in Figure~\ref{fig:pore_hbonds}
  which we used as a rough indicator of whether the head group arrangement will lead to R-double
  (Figure~\ref{fig:pore_hbonds_ft}). We see the strongest indicator of asymmetric monomer
  stacking in Pore 1 where the peaks at 0.76 $\AA^{-1}$ are clearly distinguishable, meaning
  there is periodicity at twice the $\pi$-stacking distance.

@@

>>>>

  \bibliographystyle{ieeetr}
  %\graphicspath{{./figures/supporting_figures/}}  % put all the figures here
  \graphicspath{{./figures/}}
  \maketitle

@@

  \texttt{/analysis/hexagonal\_packing.py} & \ref{M-section:rspots} & Calculate Fourier transform of
  2D array of hexagonal scatterers as shown in Figure~\ref{M-fig:hexagonal_ft} of the main text. \\ \hline

@@

  \section{Pore radius as a function of monomers per column}\label{section:pore_radius}

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{\linewidth}
  \includegraphics[width=\textwidth]{pore_radius.pdf}
  \end{subfigure}
  \caption{The pore radius increases as we increase the number of columns per pore.}\label{fig:pore_radius}
  \end{figure}

@@

  \section{Correlation Functions}\label{section:correlation_functions}

@@

  The system size, in the $z$-direction, does not significantly alter $g(z)$.
  In Figure~\ref{S-fig:correlation}, oscillations do not fully decay, so a taller
  system may be necessary to fully capture the correlation function. We therefore
  equilibrated a sandwiched system with twice as many monomers per column so that
  the system size doubled in the $z$-direction. The correlation length increases
  modestly from 4.5 $\pm$ 0.4 to 7.3 $\pm$ 1.2 \AA. However, visually, the
  correlation functions are nearly identical and oscillatory behavior persists
  throughout $g(z)$ in both cases (See Figure \ref{S-fig:z_correlation_overlay}).

  It seems likely that the discrepancy between the shape and intensity of
  R-$\pi$ generated from our simulations versus experiment is primarily due to
  highly correlated columns in our highly symmetric initial configurations.
  If the columns were uncorrelated, $g(z)$ of the ordered sandwiched system 
  would very closely resemble Figure~\ref{S-fig:z_correlation_sandwich} when we 
  average all $z$-slices of the 3D correlation function. Contributions to
  $g(z)$ from head groups in other columns within the same pore, and with head
  groups in different pores would cancel out. However, when we include all 
  $z$-slices in $g(z)$ for our system, rather than restricting it to those 
  $z$-slices within 2.1~\AA~of the center of the correlation function 
  (see Section~\ref{M-section:correlation_length}), $g(z)$ shows oscillatory 
  behavior and its correlation length nearly triples to 12.4 $\pm$ 0.7 \AA.
  (Figure~\ref{fig:z_correlation_fullbox}). The peaks of $g(z)$ are actually 
  reinforced by head groups in neighboring columns which is not expected
  in a system where columns act independently.

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{z_correlation_fullbox.pdf}
  \caption{When we average all $z$-slices of the 3D correlation function, the 
  correlation length of the ordered sandwich configuration nearly triples compared 
  to that shown in Figure~\ref{fig:z_correlation_sandwich}.}\label{fig:z_correlation_fullbox}
  \end{figure}

@@

  The appearance of R-double is likely a consequence of structure induced when
  vertically stacked monomer pairs share hydrogen bonds with a common water molecule (See
  Section~\ref{M-section:rdouble} of the main text). It is not a consequence of the 
  structure of water itself because R-double still persists if we removed water from 
  the trajectory which gives rise to the feature (Figure~\ref{fig:rdouble_water_removed}). 
  We quantified the occurence of pairing interactions by binning the $z$-positions of the center of mass of 
  the monomer head groups into 20 bins, the same number as the monomers per column. 
  This works relatively well since there is a high degree of correlation between
  columns in our system. Peaks in the distributions represent the scenario where a monomer head group
  shares a water molecule with a head group vertically above it (See Figure~\ref{fig:pore_hbonds}).
  We calculated the discrete Fourier transform of the distributions in Figure~\ref{fig:pore_hbonds}
  which we used as a rough indicator of whether the head group arrangement will lead to R-double
  (Figure~\ref{fig:pore_hbonds_ft}). We see the strongest indicator of asymmetric monomer
  stacking in Pore 1 where the peaks at 0.76 $\AA^{-1}$ are clearly distinguishable, meaning
  there is periodicity at twice the $\pi$-stacking distance.

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\textwidth]{water_removed_rdouble.png}
  \caption{When water is removed from the trajectory which produces R-double,
  R-double is still present since it is the result of uneven spacing between 
  head groups.}\label{fig:rdouble_water_removed}
  \end{figure}

@@

  \section{Influence of Cross-linking on physical properties}\label{section:xlink_physical}

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=\textwidth]{xlink_charts.pdf}
  \caption{There are minor differences between the physical properties of a
  crosslinked (xlink) system versus an uncrosslinked system. The sandwiched configuration (S)
  exhibits a smaller decrease in its ionic conductivity, while the parallel displaced (PD)
  configuration remains constant. The pore spacing of both systems decreases upon cross-linking.
  The vertical distance between stacked monomers increases in the sandwiched configuration
  and decreases in the parallel displaced configuration. The correlation length decreases
  in both configurations.}\label{fig:xlink}
  \end{figure}
